Young children rationally integrate semantic and pragmatic information sources to infer the meaning of words

### 2020-08-26 paper/manuscript.Rmd
rename model files

Experiment 3 combined the procedures from Experiment 1 and 2. It followed the same procedure as Experiment 2 but involved the same objects as Experiment 1 (Fig. 1 i-iv and Fig. S1C). In the beginning, one table was empty while there was an object (novel or familiar) on the other one. After commenting on the presence or absence of an object on each table, the speaker disappeared and a second object appeared (familiar or novel). Next, the speaker re-appeared and made the usual request. In the congruent condition, the familiar object was present in the beginning and the novel object appeared while the speaker was away (Fig. 1A and Fig. S1C - left). In this case, both the mutual exclusivity and the common ground inference pointed to the novel object as the referent. In the incongruent condition, the novel object was present in the beginning and the familiar object appeared later. In this case, the two inferences pointed to different objects (Fig. 1B and Fig. S1C - right). This resiulted in a total of 2 (congruent vs incongurent) x 12 (familiar objects) = 24 different conditions. Participants received up to 12 test trials, six in each condition, each with a different familiar and novel object. Familiar objects were the same as in Experiment 1. The positioning of the objects on the tables, the speaker and the location the speaker first turned to were counterbalanced. Participants could stop the experiment after six trials (three per condition). If a participant stopped after half of the trials, we tested an additional participant to reach a pre-registered number of data points per age group (2-, 3- and 4-year-olds).

@@

We used generalized linear mixed models to analyze the data in each of the separate experiments. These analyses were used to check whether the experimental manipulations affected children's behavior at all. Since the focus of the paper is on how information sources were integrated, we describe the details in the supplementary material.

To estimate the parameters for semantic knowledge and speaker informativeness, we adapted the rational integration to a situation in which both objects (novel and familiar) have equal prior probability (i.e., no common ground information). We used the data from Experiment 1 to then infer the parameters. That is, we inferred the intercepts and slopes for speaker informativeness (linear regression) and semantic knowledge (logistic regression) that generated model predictions to match the responses generated in Experiment 1. To estimate the parameters representing sensitivity to common ground, we used a simple logistic regression to infer which combination of intercept and slope would generate predictions that corresponded to the average proportion of correct responses measured in Experiment 2. A more detailed description of these parameters were estimated (including a graphical model) can be found in the supplementary material.

To generate model predictions we combined the parameters in the rational integration model - congruent ??

All Bayesian cognitive models were implemented in the probabilistic programming language `WebPPL` [@dippl]. The corresponding model code can be found in the associated online repository. To generate model predictions, we estimated age sensitive parameter distributions for semantic knowledge (by familiar object), speaker informativeness and sensitivity to common ground and then passed them through the model in line with the different ways in which they can be combined and aligned (Figure \@ref(fig:bda)). The resulting predictions come in the form of distributions of developmental trajectories for each familiar object in the congruent and the incongruent condition.

>>>>

Experiment 3 combined the procedures from Experiment 1 and 2. It followed the same procedure as Experiment 2 but involved the same objects as Experiment 1 (Fig. 1 i-iv and Fig. S1C). In the beginning, one table was empty while there was an object (novel or familiar) on the other one. After commenting on the presence or absence of an object on each table, the speaker disappeared and a second object appeared (familiar or novel). Next, the speaker re-appeared and made the usual request. In the congruent condition, the familiar object was present in the beginning and the novel object appeared while the speaker was away (Fig. 1A and Fig. S1C - left). In this case, both the mutual exclusivity and the common ground inference pointed to the novel object as the referent. In the incongruent condition, the novel object was present in the beginning and the familiar object appeared later. In this case, the two inferences pointed to different objects (Fig. 1B and Fig. S1C - right). This resulted in a total of 2 (congruent vs incongurent) x 12 (familiar objects) = 24 different conditions. Participants received up to 12 test trials, six in each condition, each with a different familiar and novel object. Familiar objects were the same as in Experiment 1. The positioning of the objects on the tables, the speaker and the location the speaker first turned to were counterbalanced. Participants could stop the experiment after six trials (three per condition). If a participant stopped after half of the trials, we tested an additional participant to reach a pre-registered number of data points per age group (2-, 3- and 4-year-olds).

@@

To analyze how the manipulations in each experiment affected children's behavior, we used generalized linear mixed models. Since the focus of the paper is on how information sources were integrated, we moved these models to the supplementary material. In the following, we focus on the cognitive models. A detailed, mathematical description, of the different cognitive models along with details about estimation procedures and priors can be found in the supplementary material. All Bayesian cognitive models were implemented in the probabilistic programming language `WebPPL` [@dippl]. The corresponding model code can be found in the associated online repository.

As a first step, we used the data from Experiment 1 and 2 to estimate children's developing sensitivity to each information source. To estimate the parameters for semantic knowledge ($\theta$) and speaker informativeness ($\alpha$), we adapted the rational integration model to a situation in which both objects (novel and familiar) have equal prior probability (i.e., no common ground information). We used the data from Experiment 1 to then infer the parameters. That is, we inferred the intercepts and slopes for speaker informativeness (linear regression) and semantic knowledge (logistic regression) that generated model predictions to match the responses generated in Experiment 1. To estimate the parameters representing sensitivity to common ground ($\rho$), we used a simple logistic regression to infer which combination of intercept and slope would generate predictions that corresponded to the average proportion of correct responses measured in Experiment 2. A more detailed description of these parameters were estimated (including a graphical model) can be found in the supplementary material.

To generate model predictions for the rational integration model we combined the parameters according to the model formula (Fig. 1F). As mentioned above, common ground information could either be aligned or in conflict with the other information sources. In the congruent condition, the novel object was also the discourse novel object and thus had the prior probability $\rho$. In the incongruent condition, the novel object was the "old" object and thus had the prior probability of $1 - \rho$.

### 2020-08-26 paper/supplementary_material.Rmd
advance paper reformatting
rename model files

In Experiment 1, on one table, there was a familiar object, on the other table, there was a novel object (drawn for the purpose of the study). The speaker requested an object by saying "Oh cool, there is a [non-word] on the table, how neat, can you give me the [non-word]?". Children responded by touching one of the objects. The location of the novel object (left or right table) and the animal character were counterbalanced. Each child received 12 trials, one with each familiar object. The novel object also changed from trial to trial. We coded as correct choice if children chose the novel object as the referent of the novel word.

@@

All Bayesian cognitive models were implemented in the probabilistic programming language `WebPPL` [@dippl]. The corresponding model code can be found in the associated online repository (file `webppl/prediction_model.wppl`). To generate model predictions, we estimated age sensitive parameter distributions for semantic knowledge (by familiar object), speaker informativeness and sensitivity to common ground and then passed them through the model in line with the different ways in which they can be combined and aligned (Figure \@ref(fig:bda)). The resulting predictions come in the form of distributions of developmental trajectories for each familiar object in the congruent and the incongruent condition.

@@

To estimate $\phi$ as well as the other parameters, we make use of all the available data. The model code can be found in the associated online repository (file: `webppl/explanation_mixture_model.wppl`). The posterior distribution for the bias parameter $\phi$ is shown in figure \@ref(fig:fig10)A. It suggests that the mutual exclusivity inference is weighted as slightly more important compared to the common ground inference.

@@

The model code can be found in the associated online repository (file: `webppl/explanation_mixture_model.wppl`). Figure \@ref(fig:fig10)B visualizes the developmental trajectory of the bias parameter. Based on this model, the common ground inference seems to decrease in importance compared to the mutual exclusivity inference with age.

>>>>

In Experiment 1, on one table, there was a familiar object, on the other table, there was a novel object (drawn for the purpose of the study). The speaker requested an object by saying "Oh cool, there is a [non-word] on the table, how neat, can you give me the [non-word]?". Children responded by touching one of the objects. The location of the novel object (left or right table) and the animal character were counterbalanced. We coded as correct choice if children chose the novel object as the referent of the novel word.

@@

All Bayesian cognitive models were implemented in the probabilistic programming language `WebPPL` [@dippl]. The corresponding model code can be found in the associated online repository (file `webppl/prediction_all_models.wppl`). To generate model predictions, we estimated age sensitive parameter distributions for semantic knowledge (by familiar object), speaker informativeness and sensitivity to common ground and then passed them through the model in line with the different ways in which they can be combined and aligned (Figure \@ref(fig:bda)). The resulting predictions come in the form of distributions of developmental trajectories for each familiar object in the congruent and the incongruent condition.

@@

To estimate $\phi$ as well as the other parameters, we make use of all the available data. The model code can be found in the associated online repository (file: `webppl/explanation_bias_model.wppl`). The posterior distribution for the bias parameter $\phi$ is shown in figure \@ref(fig:fig10)A. It suggests that the mutual exclusivity inference is weighted as slightly more important compared to the common ground inference.

@@

The model code can be found in the associated online repository (file: `webppl/explanation_dev_bias_model.wppl`). Figure \@ref(fig:fig10)B visualizes the developmental trajectory of the bias parameter. Based on this model, the common ground inference seems to decrease in importance compared to the mutual exclusivity inference with age.
